---
title: "INF 550 Intro to NEON Exercises Part 1"
author: "Natasha Wesely"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(neonUtilities)
library(geoNEON)
library(sp)
```

# 2.9.1.1 Part 1: Sign up for and Use an NEON API Token:

```{r}
# call in your unique NEON API token
load('neon_token_source.Rdata')

# demonstrate the you can use your NEON API token to pull some data
veglist <- loadByProduct(dpID="DP1.10098.001", site="WREF", package="basic", check.size=FALSE, token = NEON_TOKEN)

# prep the data further
vegmap <- getLocTOS(veglist$vst_mappingandtagging, 
                    "vst_mappingandtagging")
veg <- merge(veglist$vst_apparentindividual, vegmap, 
             by=c("individualID","namedLocation",
                  "domainID","siteID","plotID"))
```



# 2.9.1.2 Part 2: Further Investigation of NEON TOS Vegetation Structure Data
## Question 1
Convert the previous diagram plot into a ggplot.

```{r}

veg %>% 
  filter(plotID == "WREF_075") %>%
  ggplot() +
  geom_point(aes(x = adjEasting, y = adjNorthing, 
                 size = stemDiameter), shape = 1) +
  labs(x = "Easting", y = "Northing")
```

## Question 2
Set the color your circles to be a function of each species.

```{r}
veg %>% 
  filter(plotID == "WREF_075") %>%
  ggplot() +
  geom_point(aes(x = adjEasting, y = adjNorthing, 
                 size = stemDiameter, color = taxonID), shape = 1) +
  labs(x = "Easting", y = "Northing")
```


## Question 3
Generate a histogram of tree heights for each plot. Color your stacked bar as a function of each species.

```{r}
veg %>% ggplot() +
  geom_histogram(aes(x = height, fill = taxonID)) +
  facet_wrap(~ plotID)
```

## Question 4
Use dplyr to remove dead trees.
```{r}
# what are all the options in plantStatus?
unique(veg$plantStatus)

# filter out any dead/downed trees
vegLiv = veg %>%
  filter(
    plantStatus != "Dead, broken bole", 
    plantStatus != "Lost, fate unknown", 
    plantStatus != "Standing dead", 
    plantStatus != "No longer qualifies", 
    plantStatus != "Downed")

# make sure your filtering worked
unique(vegLiv$plantStatus)
```

## Question 5
Create a simple linear model that uses Diameter at Breast Height (DBH) and height to predict allometries. Print the summary information of your model.

```{r}
model = lm(baseCrownHeight ~ stemDiameter + height,
            data = vegLiv)
summary(model)
```

## Question 6
Plot your linear model.
```{r}
plot.df = vegLiv %>%
  select(stemDiameter, height, baseCrownHeight, taxonID) %>%
  na.omit()

plot.df$predictions = predict(model)

plot.df %>% ggplot() +
  geom_point(aes(x = baseCrownHeight, y = predictions, color = taxonID)) +
  geom_smooth(aes(x = baseCrownHeight, y = predictions), method = "lm") +
  labs(y = "Predicted Base Crown Height", x = "Observed Base Crown")
```

I realize I did not make separate lines for the different species, but that is because I only have 8 observations (2 observations for VAPA, 2 observations for MEFE, and 3 observations for COCOC). Having separate lines would be nonsensical. 

## Question 7
Answer the following questions.

**What do you think about your simile linear model? What are its limitations?**

My simple linear model is not great but not terrible. The biggest issue with this model that makes this model questionable is the extremely small sample size. Linear models have a lot of limitations, including the assumption of normally distributed residuals, the assumption of a linear relationship between your response and predictor(s), and the assumption of homoscedasticty. It's very challenging characterize complicated relationships with a simple linear model in general.

**How many unique species are present at WREF?**
```{r}
length(unique(veg$taxonID))
```

There are 25 unique species at the WREF site.

**What are the top_5 trees based on height? Diameter?**
```{r}
# What are the top 5 trees based on height?
vegHT = vegLiv %>% arrange(desc(height)) %>%
  select(individualID, height)

head(vegHT, n = 5)

# what are the top 5 trees based on diameter?
vegDM = vegLiv %>% arrange(desc(stemDiameter)) %>%
  select(individualID, stemDiameter)

head(vegDM, n = 5)
```


**What proportion of sampled trees are dead?**
```{r}
liveTrees = length(unique(vegLiv$individualID))
allTrees = length(unique(veg$individualID))
livePropor = liveTrees / allTrees
(deadTreesProportion = 1 - livePropor)
```



