---
title: "INF 550 Section 4.13"
author: "Natasha Wesely"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4.13 Digital Repeat Photography Coding Lab

## 4.13.1 Quantifying haze and redness to evaluate California wildfires

### Question 1
**Pull mid-day imagery for September 1-7th, 2019 and 2020 for the canopy-level camera NEON.D17.SOAP.DP1.00033. Create a 2-panel plot showing those images in 2019 (left) and 2020 (right).**

```{r}

```


### Question 2
**Use the hazeR package to quantify the haze in each of those images. Print a summary of your results.**

### Question 3
**Generate a density function RGB plot for your haziest image in 2020, and one for the same date in 2019. Create a 2-panel plot showing 2019 on the left and 2020 on the right.**

### Question 4
**Pull timesseries data via the phenocamapi package. Calculate the difference in the rcc90 between 2019 and 2020 over the same time period as your images.**

### Question 5
**Create a summary plot showing haze as a bar and the differenece in rcc90 from question 4 as a timersies.**

### Question 6
**Answer the following questions:**

**Does the hazeR package pick up smokey images?**

**If you were to use color coordinates, which color band would be most useful to highlight smoke and why?**

