---
title: "INF 550 Section 6.15"
author: "Natasha Wesely"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lidR)
library(neonUtilities)
load("neon_token_source.Rdata")

```

# NEON AOP Coding Lab
## Question 1
**1: How are these two forests similar? How are they different? (3-5 sentences)**

The forest at GUAN is a tropical dry forest. It's mixed forest type with deciduous, evergreen, and scrub vegetation. It has a warm and wet climate compared to BART.

The forest at the BART site is a mixed temperate forest. This site had been heavily disturbed by logging activities in the past. BART is a much cooler site than GUAN.

## Question 2
**2: Using this NEON tutorial and the tutorials we’ve covered in this textbook (hint: you did half of this workflow in your very first coding lab) pull the NEON AOP derived Canopy Height Model (CHM, DP3.30015.001) and High-resolution orthorectified camera imagery mosaic DP3.30010.001 for each forest and overlay the NEON TOS Woody Vegetation Structure DP1.10098.001 basal diameter data product to evaluate how well AOP captures trees in each forest.**

```{r}
# pull NEON AOP Canopy Height Model 
chm_dp = as.character('DP1.10098.001')
# for BART
chm_bart = loadByProduct(dpID = chm_dp, 
                     site="BART",
                     token=NEON_TOKEN,
                     # year = 2020,
                     startdate="2022-01", enddate="2022-12",
                     package="basic",
                     check.size=F)

zipsByProduct(dpID=chm_dp,
              site = "BART",
              startdate = "2022-01",
              enddate = "2022-12",
              package = "basic",
              check.size = FALSE,
              savepath = "./",
              token=NEON_TOKEN)

# for GUAN
chm_guan = byFileAOP(dpID = "DP3.30015.001", 
                     site="GUAN",
                     token=NEON_TOKEN,
                     year = 2020,
                     # startdate="2022-05", enddate="2022-05",
                     # package="basic",
                     check.size=F)

# pull NEON AOP high res orthorectified imagery
# for BART
# for GUAN

# pull NEON TOS Woody Veg Structure 
# for BART
wood_bart = loadByProduct(
  dpID = as.character("DP1.10098.001"),
  site = "BART",
   token=NEON_TOKEN,
  startdate="2022-01", 
  enddate="2022-12",
  package="basic",
  check.size=F
)
# for GUAN


```


**Generate a labeled 2x2 plot panel including:**

**Each RGB image with basal diameter overlaid**

**Each CHM with basal diameter overlaid**

**Write 2-3 sentences summarizing your findings and thoughts.**

## Question 3
**Use the byTileAOP function of the neonUtilities package to pull a subset of the descrete LiDAR pointcloud for each forest (Hint: You can feed byTileAOP Easting and Northing from your Vegetation Structure dataframe(s)). Use the structural_diversity_metrics function that you defined in section 6.5 of the textbook to process discrete return LiDAR for each site and generte structural diversity metrics.

Using lidR generate a labeled 2-panel plot of your canopy height model for each forest
Using lidR generate a labeled 2-panel plot of a cross-section for each forest
Use section 6.5.3 Comparing Metrics Between Forests to compare each forest and generate a a clean summary table via kable
Using Table 2 from LaRue, Wagner, et al. (2020) as a reference, write 1-2 paragraphs summarizing the differences in forest structural diversity between the two forests and how they may relate to your answers to Question 1.**
