---
title: "INF 550 Section 4.13"
author: "Natasha Wesely"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# devtools::install_github("https://github.com/bnasr/phenocamapi")
# utils::install.packages('xROI', repos = "http://cran.us.r-project.org" )
library(data.table)
library(phenocamapi)
library(lubridate)
library(jpeg)
library(hazer)
library(dplyr)
library(tidyverse)
```


# 4.13 Digital Repeat Photography Coding Lab

## 4.13.1 Quantifying haze and redness to evaluate California wildfires

## Question 1
**Pull mid-day imagery for September 1-7th, 2019 and 2020 for the canopy-level camera NEON.D17.SOAP.DP1.00033. Create a 2-panel plot showing those images in 2019 (left) and 2020 (right).**

```{r}
# open a temporary directory
jpeg_dir <- "data_4_13"

# download a subset
download_midday_images(site = 'NEON.D17.SOAP.DP1.00033', # which site
                       y = 2019:2020, # which year(s)
                       months = 9, # which month(s)
                       days = 1:7, # which days on month(s)
                       download_dir = jpeg_dir) # where on your computer

# list of paths to the downloaded files
midday_paths <- dir(jpeg_dir, pattern = '.jpg*', full.names = TRUE)

# set up the plot dimensions
par(mar= c(0,0,0,0), 
    mfrow=c(3,2), 
    oma=c(0,0,3,0)
    )

# make a vector of the dates just to make the graphic nice
dates = c("9/2/2019","9/4/2019","9/6/2019","9/2/2020","9/4/2020","9/6/2020")

for (i in 1:length(midday_paths)) {
  img <- readJPEG(midday_paths[i])
  plot(0:1,0:1, type='n', axes= FALSE, xlab= '', ylab = '')
  rasterImage(img, 0, 0, 1, 1)
  mtext(dates[i], line = -2)
}

```


## Question 2
**Use the hazeR package to quantify the haze in each of those images. Print a summary of your results.**

```{r}

# create an empty matrix to fill with haze and A0 values
haze_mat <- data.frame()

# loop through each image and calc the haze factor
for (i in 1:length(midday_paths)){
  image_path <- midday_paths[i]
  img <- jpeg::readJPEG(image_path)
  # calc the haze level
  haze_degree <- getHazeFactor(img)
  # store the overall haze level for the image in the matix
  haze_mat <- rbind(haze_mat,
                    data.frame(file = as.character(image_path),
                               haze = haze_degree[1],
                               A0 = haze_degree[2]))
  
}

# Print a summary of your results
head(haze_mat)
summary(haze_mat)
```


## Question 3
**Generate a density function RGB plot for your haziest image in 2020, and one for the same date in 2019. Create a 2-panel plot showing 2019 on the left and 2020 on the right.**
```{r}
# split the matrix by year
haze_mat2020 = haze_mat[which(grepl(pattern = "2020", x = haze_mat$file) == TRUE),]
haze_mat2019 = haze_mat[which(grepl(pattern = "2019", x = haze_mat$file) == TRUE),]

# which image is the most hazy in 2020? - Sept 2
mostHazy2020 = haze_mat2020[which(haze_mat2020$haze == max(haze_mat2020$haze)),]

# find the same day in 2019
match2019 = haze_mat2019[which(grepl(pattern = "09_02", x = haze_mat2019$file) == TRUE),]

# read the jpgs
rgb_array1 = jpeg::readJPEG(mostHazy2020[1,1])
rgb_array2 = jpeg::readJPEG(match2019[1,1])

# color channels extracted from the matrix
red_vector1 <- rgb_array1[,,1]
green_vector1 <- rgb_array1[,,2]
blue_vector1 <- rgb_array1[,,3]
red_vector2 <- rgb_array2[,,1]
green_vector2 <- rgb_array2[,,2]
blue_vector2 <- rgb_array2[,,3]

# plotting
par(mar=c(5,4,4,2), mfrow=c(1,2))
plot(density(red_vector2), col = 'red', lwd = 2,
     main = 'Matching Image in 2019', ylim = c(0,5))
lines(density(green_vector2), col = 'green', lwd = 2)
lines(density(blue_vector2), col = 'blue', lwd = 2)
plot(density(red_vector1), col = 'red', lwd = 2,
     main = 'Haziest Image in 2020', ylim = c(0,5))
lines(density(green_vector1), col = 'green', lwd = 2)
lines(density(blue_vector1), col = 'blue', lwd = 2)

```


## Question 4
**Pull timesseries data via the phenocamapi package. Calculate the difference in the rcc90 between 2019 and 2020 over the same time period as your images.**

```{r}
# define the date ranges
dateSeq1 = seq(ymd('2019-09-01'),ymd('2019-09-07'), by = '1 day')
dateSeq2 = seq(ymd('2020-09-01'),ymd('2020-09-07'), by = '1 day')
dateSeq = c(dateSeq1, dateSeq2)

# grab the data 
data = read_csv("data_4_13/NEON.D17.SOAP.DP1.00033_EN_1000_1day.csv", skip = 24) %>%
  filter(date %in% dateSeq) %>% 
  select(date, year, doy, rcc_90) %>%
  mutate(DayMonth = format(date, "%m-%d"))

(RccDiffs = data %>%
  group_by(DayMonth) %>%
  summarise(Rcc_diff = diff(rcc_90)) %>%
    ungroup())
```

The mean Rcc (90th percentile) in 2020 is `r mean(RccDiffs$Rcc_diff)` higher than the mean Rcc (90th percentile) in 2019.


## Question 5
**Create a summary plot showing haze as a bar and the difference in rcc90 from question 4 as a timerseries.**

```{r}
ggplot() +
  geom_point(data = RccDiffs, aes(x = DayMonth, y = Rcc_diff), color = "red") +
  labs(y = "Difference in Rcc between 2019 and 2020 on the Same Day")
```



## Question 6
**Answer the following questions:**

**Does the hazeR package pick up smokey images?**

No, hazeR does not seem to pick up smokey images well.

**If you were to use color coordinates, which color band would be most useful to highlight smoke and why?**

The red band would be the most useful for detecting smoke in a photo because the smokey pixels have more red in them.
