---
title: "INF 550 Section 6.15"
author: "Natasha Wesely"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(neonUtilities)
library(BiocManager)
library(rhdf5)
library(raster)
library(rgdal)
library(geoNEON)
load("neon_token_source.Rdata")

```

# NEON AOP Coding Lab
## Question 1
**1: How are these two forests similar? How are they different? (3-5 sentences)**

The forest at GUAN is a tropical dry forest. It's mixed forest type with deciduous, evergreen, and scrub vegetation. It has a warm and wet climate compared to BART.

The forest at the BART site is a mixed temperate forest. This site had been heavily disturbed by logging activities in the past. BART is a much cooler site than GUAN.

## Question 2
**2: Using this NEON tutorial and the tutorials we’ve covered in this textbook (hint: you did half of this workflow in your very first coding lab) pull the NEON AOP derived Canopy Height Model (CHM, DP3.30015.001) and High-resolution orthorectified camera imagery mosaic DP3.30010.001 for each forest and overlay the NEON TOS Woody Vegetation Structure DP1.10098.001 basal diameter data product to evaluate how well AOP captures trees in each forest.**

```{r}
# # pull NEON AOP Canopy Height Model 
# chm_dp = as.character('DP1.10098.001')
# # for BART
# chm_bart = loadByProduct(dpID = chm_dp, 
#                          site="BART",
#                          token=NEON_TOKEN,
#                          # year = 2020,
#                          startdate="2022-01", enddate="2022-12",
#                          package="basic",
#                          check.size=F)
# 
# zipsByProduct(dpID=chm_dp,
#               site = "BART",
#               startdate = "2022-01",
#               enddate = "2022-12",
#               package = "basic",
#               check.size = FALSE,
#               savepath = "./",
#               token=NEON_TOKEN)
# 
# # for GUAN
# chm_guan = byFileAOP(dpID = "DP3.30015.001", 
#                      site="GUAN",
#                      token=NEON_TOKEN,
#                      year = 2020,
#                      # startdate="2022-05", enddate="2022-05",
#                      # package="basic",
#                      check.size=F)
# 
# # pull NEON AOP high res orthorectified imagery
# # for BART
# # for GUAN
# 
# # pull NEON TOS Woody Veg Structure 
# # for BART
# wood_bart = loadByProduct(
#   dpID = as.character("DP1.10098.001"),
#   site = "BART",
#   token=NEON_TOKEN,
#   startdate="2022-01", 
#   enddate="2022-12",
#   package="basic",
#   check.size=F
# )
# # for GUAN


```


```{r, cache = T}
# make vector of sites
sites <- c("BART","GUAN")

# Download veg structure 
veglist <- loadByProduct(dpID="DP1.10098.001", 
                         site= sites,
                         package="expanded",
                         check.size=FALSE,
                         token = NEON_TOKEN)

# Unlist veg data frames
list2env(veglist,.GlobalEnv)

# get the veg map object
vegmap <- getLocTOS(veglist$vst_mappingandtagging, "vst_mappingandtagging")

# Merge
veg <- merge(veglist$vst_apparentindividual,vegmap, by=c("individualID","namedLocation","domainID","siteID","plotID"))

# Separate out the sites
vegbart <- veg %>%
  filter(plotID == "BART_013") %>%
  drop_na(adjNorthing, adjEasting)
vegguan <- veg %>%
  filter(plotID == "GUAN_043") %>%
  drop_na(adjNorthing, adjEasting)

```


```{r, cache=TRUE}
# year we want data for
yoi <- "2018"

# Download Canopy Height Model from NEON API
# BART
byTileAOP(dpID="DP3.30015.001", # CHM
          site= "BART",
          year= yoi,
          easting = vegbart$adjEasting,
          northing = vegbart$adjNorthing,
          check.size=FALSE,
          token = NEON_TOKEN)
# GUAN
byTileAOP(dpID="DP3.30015.001", # CHM
          site= "GUAN",
          year= yoi,
          easting = vegguan$adjEasting,
          northing = vegguan$adjNorthing,
          check.size=FALSE,
          token = NEON_TOKEN)

# Download High-resolution orthorectified camera imagery mosaic from NEON API
# BART
byTileAOP(dpID="DP3.30010.001", 
          site="BART",
          year = yoi,
          easting = vegbart$adjEasting,
          northing = vegbart$adjNorthing,
          check.size=FALSE,
          token = NEON_TOKEN)
# GUAN
byTileAOP(dpID="DP3.30010.001",
          site="GUAN",
          year = yoi,
          easting = vegguan$adjEasting,
          northing = vegguan$adjNorthing,
          check.size=FALSE,
          token = NEON_TOKEN)

# convert to rasters
chmbart <- raster('./DP3.30015.001/neon-aop-products/2018/FullSite/D01/2018_BART_4/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D01_BART_DP3_317000_4879000_CHM.tif')
chmguan <- raster('./DP3.30015.001/neon-aop-products/2018/FullSite/D04/2018_GUAN_1/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D04_GUAN_DP3_725000_1988000_CHM.tif')

```

**Generate a labeled 2x2 plot panel including:**

**Each RGB image with basal diameter overlaid**

**Each CHM with basal diameter overlaid**

```{r, cache=TRUE}
# PLOTTING

# make image bricks
imagebart <-brick('./DP3.30010.001/neon-aop-products/2018/FullSite/D01/2018_BART_4/L3/Camera/Mosaic/2018_BART_4_317000_4879000_image.tif')
imageguan <- brick('./DP3.30010.001/neon-aop-products/2018/FullSite/D04/2018_GUAN_1/L3/Camera/Mosaic/2018_GUAN_1_725000_1988000_image.tif')

# the arrangement of the subplots
par(mfrow=c(2,2))
# plot the canopy height model for BART
plot(chmbart, xlim=c(317500,317540), ylim=c(4879020,4879050), main="BART CHM", xlab = "Easting", ylab = "Northing")
# add the woody veg structure for BART
points(vegbart$adjEasting,vegbart$adjNorthing, pch=1, cex=vegbart$stemDiameter/100/2, col = "red")
# plot the canopy height model for GUAN
plot(chmguan, xlim=c(725780,725850), ylim=c(1988000,1988100), main="GUAN CHM", xlab = "Easting", ylab = "Northing")
# add the woody veg str for GUAN
points(vegguan$adjEasting,vegguan$adjNorthing, pch=1, cex=vegguan$stemDiameter/100/2, col = "red")
# plot the ortho image for BART
plot.new()
plot.window(xlim=c(317500,317540), ylim=c(4879020,4879050))
plotRGB(imagebart, r=1,g=2,b=3, axes = TRUE, main="BART Image", add=TRUE)
# add the woody veg structure for BART
points(vegbart$adjEasting,vegbart$adjNorthing, pch=1, cex=vegbart$stemDiameter/100/2, col = "red")
# plot the orthose image for GUAN
plot.new()
plot.window(xlim=c(725780,725850), ylim=c(1988000,1988100))
plotRGB(imageguan, r=1,g=2,b=3, axes = TRUE, main="GUAN Image", add=TRUE)
# add the woody veg str for GUAN
points(vegguan$adjEasting,vegguan$adjNorthing, pch=1, cex=vegguan$stemDiameter/100/2, col = "red")

```


**Write 2-3 sentences summarizing your findings and thoughts.**

The NEON AOP data is quite course, which makes identifying individual trees within a canopy nearly impossible. However the combination of the AOP data with the individual tree measurements from the woody vegetation data set could enhance the AOP data. 

## Question 3
**Use the byTileAOP function of the neonUtilities package to pull a subset of the descrete LiDAR pointcloud for each forest (Hint: You can feed byTileAOP Easting and Northing from your Vegetation Structure dataframe(s)). Use the structural_diversity_metrics function that you defined in section 6.5 of the textbook to process discrete return LiDAR for each site and generate structural diversity metrics.**

```{r, cache=TRUE}

library(lidR, quietly = T)
library(gstat, quietly = T)

# get the lidar pointcloud for BART
byTileAOP(dpID="DP1.30003.001",
          site= "BART",
          year= yoi,
          easting = extent(chmbart)[1],
          northing = extent(chmbart)[3],
          check.size=FALSE,
          token = NEON_TOKEN)

# get the lidar pointcloud for GUAN
byTileAOP(dpID="DP1.30003.001",
          site= "GUAN",
          year= yoi,
          easting = extent(chmguan)[1],
          northing = extent(chmguan)[3],
          check.size=FALSE,
          token = NEON_TOKEN)

# convert the raw data to an actual pointcloud 
pcbart <- lidR::readLAS('/Users/natashawesely/Documents/GitHub/INF550/DP1.30003.001/neon-aop-products/2018/FullSite/D01/2018_BART_4/L1/DiscreteLidar/ClassifiedPointCloud/NEON_D01_BART_DP1_317000_4879000_classified_point_cloud_colorized.laz')
pcguan <- lidR::readLAS('/Users/natashawesely/Documents/GitHub/INF550/DP1.30003.001/neon-aop-products/2018/FullSite/D04/2018_GUAN_1/L1/DiscreteLidar/ClassifiedPointCloud/NEON_D04_GUAN_DP1_725000_1988000_classified_point_cloud_colorized.laz')

```


**Using lidR generate a labeled 2-panel plot of your canopy height model for each forest**

```{r, eval=FALSE}
lidR::plot(pcbart)
lidR::plot(pcguan)
```


```{r, echo=FALSE}
library(png)
pngs = c("BARTchm.png","GUANchm.png")
labs = c("BART", "GUAN")

par(mfrow=c(1,2))

for (i in 1:length(pngs)) {
  img <- readPNG(pngs[i])
  plot(0:1,0:1, type='n', axes= FALSE, xlab= '', ylab = '') 
  rasterImage(img, 0, 0, 1, 1)
  mtext(labs[i], line = -2, col = "red")
}
```

**Using lidR generate a labeled 2-panel plot of a cross-section for each forest**

```{r, cache = T}
## BART
# Correct for elevation
# Choose a 40 x 40 m spatial extent, which is the extent for NEON base plots.
# First set the center of where you want the plot to be
xbart <- 317500 #easting
ybart <- 4879500 #northing
#Cut out a 200 x 200 m buffer by adding 100 m to easting and
#northing coordinates (x,y).
data.200m.b <-
  clip_rectangle(pcbart,
                 xleft = (xbart - 100), ybottom = (ybart - 100),
                 xright = (xbart + 100), ytop = (ybart + 100))
# Correct for ground height using a kriging function to interpolate
#elevation from ground points in the .laz file.
dtm.b <- grid_terrain(data.200m.b, 1, kriging(k = 10L))
data.200m.b <- normalize_height(data.200m.b, dtm.b)
#Will often give a warning if not all points could be corrected,
#but visually check to see if it corrected for ground height.
lidR::plot(data.200m.b)
#There's a few uncorrected points
#Clip 20 m out from each side of the easting and northing
#coordinates (x,y).
data.40m.b <-
  clip_rectangle(data.200m.b,
                 xleft = (xbart - 20), ybottom = (ybart - 20),
                 xright = (xbart + 20), ytop = (ybart + 20))

data.40m.b@data$Z[data.40m.b@data$Z <= .5] <- NA
#This line filters out all z_vals below .5 m as we are less interested in shrubs/trees.
# visualize the clipped plot point cloud
lidR::plot(data.40m.b)

## GUAN
# Correct for elevation
# Choose a 40 x 40 m spatial extent, which is the extent for NEON base plots.
# First set the center of where you want the plot to be
xguan <- 725500 #easting
yguan <- 1988500 #northing
# Cut out a 200 x 200 m buffer by adding 100 m to easting and
#northing coordinates (x,y).
data.200m.g <-
  clip_rectangle(pcguan,
                   xleft = (xguan - 100), ybottom = (yguan - 100),
                   xright = (xguan + 100), ytop = (yguan + 100))
# Correct for ground height using a kriging function to interpolate
#elevation from ground points in the .laz file.
dtm.g <- grid_terrain(data.200m.g, 1, kriging(k = 10L))
## Warning: There were 23 degenerated ground points. Some X Y coordinates were
## repeated but with different Z coordinates. min Z were retained.
data.200m.g <- normalize_height(data.200m.g, dtm.g)
#Will often give a warning if not all points could be corrected,
#but visually check to see if it corrected for ground height.
lidR::plot(data.200m.g)
#There's a few uncorrected points
#Clip 20 m out from each side of the easting and northing
#coordinates (x,y).
data.40m.g <-
  clip_rectangle(data.200m.g,
                   xleft = (xguan - 20), ybottom = (yguan - 20),
                   xright = (xguan + 20), ytop = (yguan + 20))

data.40m.g@data$Z[data.40m.g@data$Z <= .5] <- NA
# This line filters out all z_vals below .5 m
# visualize the clipped plot point cloud
lidR::plot(data.40m.g)

```

```{r, echo=F}
library(png)
# BARTpng = readPNG("barlettPointCloud.png")
# GUANpng = readPNG("GUANpointCloud.png")
pngs = c("barlettPointCloud.png","GUANpointCloud.png")
labs = c("BART", "GUAN")

par(mfrow=c(1,2))

for (i in 1:length(pngs)) {
  img <- readPNG(pngs[i])
  plot(0:1,0:1, type='n', axes= FALSE, xlab= '', ylab = '') 
  rasterImage(img, 0, 0, 1, 1)
  mtext(labs[i], line = -2, col = "red")
}
```



**Use section 6.5.3 Comparing Metrics Between Forests to compare each forest and generate a a clean summary table**

```{r, cache = T}
# write function to run all 13 veg structure metrics in a single function.
structural_diversity_metrics <- function(data.40m,x,y) {
  chm <- grid_canopy(data.40m, res = 1, dsmtin())
  mean.max.canopy.ht <- mean(chm@data@values, na.rm = TRUE)
  max.canopy.ht <- max(chm@data@values, na.rm=TRUE)
  rumple <- rumple_index(chm) #calculate rumple, a ratio of outer canopy surface area to ground surface area (1600 m^2)
  top.rugosity <- sd(chm@data@values, na.rm = TRUE) #top rugosity, the standard deviation of pixel values in chm = measure of outer canopy roughness
  cells <- length(chm@data@values)
  chm.0 <- chm
  chm.0[is.na(chm.0)] <- 0
  zeros <- which(chm.0@data@values == 0)
  deepgaps <- length(zeros) #number of deep gaps
  deepgap.fraction <- deepgaps/cells #deep gap fraction, the number of deep gaps in the chm relative to total number of chm pixel
  cover.fraction <- 1 - deepgap.fraction #cover fraction, the inverse of deep gap fraction
  vert.sd <- cloud_metrics(data.40m, sd(Z, na.rm = TRUE)) #height SD, the standard deviation of height values for all points
  
  17
  
  sd.1m2 <- grid_metrics(data.40m, sd(Z), 1)
  sd.sd <- sd(sd.1m2[,3], na.rm = TRUE) # The standard deviation of height values
  Zs <- data.40m@data$Z
  Zs <- Zs[!is.na(Zs)]
  entro <- entropy(Zs, by = 1) # Quantifies diversity & evenness of point cloud heights
  gap_frac <- gap_fraction_profile(Zs, dz = 1, z0=3)
  GFP.AOP <- mean(gap_frac$gf) # Gap fraction profile, assesses the distribution of gaps in the canopy volume
  LADen<-LAD(Zs, dz = 1, k=0.5, z0=3)
  VAI.AOP <- sum(LADen$lad, na.rm=TRUE) # Leaf area density
  VCI.AOP <- VCI(Zs, by = 1, zmax=100) # A vertical complexity index
  out.plot <- data.frame(
    matrix(c(x, y, mean.max.canopy.ht,max.canopy.ht,
             rumple,deepgaps, deepgap.fraction,
             cover.fraction, top.rugosity, vert.sd,
             sd.sd, entro, GFP.AOP, VAI.AOP,VCI.AOP),
           ncol = 15))
  colnames(out.plot) <-
    c("easting", "northing", "mean.max.canopy.ht.aop",
      "max.canopy.ht.aop", "rumple.aop", "deepgaps.aop",
      "deepgap.fraction.aop", "cover.fraction.aop",
      "top.rugosity.aop","vert.sd.aop","sd.sd.aop",
      "entropy.aop", "GFP.AOP.aop",
      "VAI.AOP.aop", "VCI.AOP.aop")
  print(out.plot)
}

# run the function for each site
BART_structural_diversity <- structural_diversity_metrics(data.40m.b,xbart,ybart)
GUAN_structural_diversity <- structural_diversity_metrics(data.40m.g,xguan,yguan)

# combine the function runs for each site
combined_results=rbind(BART_structural_diversity,
                       GUAN_structural_diversity)

# Add row names for clarity
row.names(combined_results)=c("BART","GUAN")
# print the comparison 
combined_results

```


**Using Table 2 from LaRue, Wagner, et al. (2020) as a reference, write 1-2 paragraphs summarizing the differences in forest structural diversity between the two forests and how they may relate to your answers to Question 1.**

Unsurprisingly, these forested sites are quite different. BART is on average taller with a higher leaf area density compared to GUAN. BART also has a higher standard deviation for height measurements, indicating that the forest canopy is more vertically diverse than GUAN. This higher height standard deviation may also indicate more different species close together compared to GUAN. Perhaps because the tree/leaf density is less intense at GUAN, the trees don't have to compete for light as much, therefore there is more canopy height homogeneity and less species diversity. 

